---
title: "Models With Images vs Without Images"
author: "Anonymous"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(broom)
library(rstatix)
library(ggrepel)
```

## 1. Load Data

```{r load-data}
file_images <- "../output_csvs/image_comparison_grade_dist.csv"  # edit path if needed

raw <- readr::read_csv(file_images, show_col_types = FALSE)
glimpse(raw)
```

Assumptions (from the file):

* Column 1 (`Unnamed: 0`) is an index → we drop it.
* Remaining columns are model score distributions:

  * Plain models: e.g. `gpt4.1`, `o3`, `gemini_2.0_flash`
  * Image variants: e.g. `gpt4.1_images`, `o3_images`, `gemini_2.0_flash_images`

```{r clean-data}
# Drop index column
dat <- raw[, -1]

names(dat)
```

## 2. Identify Image vs Non-Image Model Pairs

We assume image variants are named `*_images`.

```{r identify-pairs}
# Columns that are image variants
image_cols <- names(dat)[str_ends(names(dat), "_images")]
image_cols

# Base model names (without _images)
base_models <- sub("_images$", "", image_cols)
base_models

# Check that each base model exists as a non-image column
non_image_cols <- base_models[base_models %in% names(dat)]
non_image_cols

# Keep only pairs where both exist
pairs <- tibble(
  base_model      = base_models,
  image_model_col = image_cols
) %>%
  filter(base_model %in% names(dat))

pairs
```

At this point, for your file, `pairs` should look roughly like:

* `gpt4.1` ↔ `gpt4.1_images`
* `o3` ↔ `o3_images`
* `gemini_2.0_flash` ↔ `gemini_2.0_flash_images`

## 3. Long-Format Data for Plotting

```{r long-format}
# Build a long data frame: model, variant (with/without images), score
long_list <- lapply(seq_len(nrow(pairs)), function(i) {
  base  <- pairs$base_model[i]
  img   <- pairs$image_model_col[i]

  tibble(
    model_base = base,
    variant    = c(
      rep("no_images", nrow(dat)),
      rep("with_images", nrow(dat))
    ),
    score      = c(dat[[base]], dat[[img]])
  )
})

scores_long <- bind_rows(long_list)

scores_long
```

Quick summary:

```{r descriptives}
summary_stats <- scores_long %>%
  group_by(model_base, variant) %>%
  summarise(
    n      = n(),
    mean   = mean(score, na.rm = TRUE),
    median = median(score, na.rm = TRUE),
    sd     = sd(score, na.rm = TRUE),
    .groups = "drop"
  )

summary_stats
```

## 4. Visual Comparison

### 4.1 Boxplots per Model (with vs without images)

```{r boxplots, fig.width=10, fig.height=0.4 * n_distinct(scores_long$model_base) + 4}
scores_long %>%
  ggplot(aes(x = variant, y = score, fill = variant)) +
  geom_boxplot(outlier.alpha = 0.4) +
  facet_wrap(~ model_base, scales = "free_y") +
  labs(
    title = "Model Scores: With Images vs Without Images",
    x = "Variant",
    y = "Score"
  ) +
  theme_minimal()
```

### 4.2 Density Plot (stacked by model/variant)

```{r density, fig.width=10, fig.height=0.4 * n_distinct(scores_long$model_base) + 4}
scores_long %>%
  ggplot(aes(x = score, fill = variant)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~ model_base, scales = "free") +
  labs(
    title = "Score Distributions by Model and Image Variant",
    x = "Score",
    y = "Density"
  ) +
  theme_minimal()
```

## 5. Mann–Whitney (Wilcoxon Rank-Sum) Tests

We treat the with-images and no-images distributions as **independent samples** and compare them using a Mann–Whitney U / Wilcoxon rank-sum test for each base model.

```{r mann-whitney}
mw_results <- purrr::map_df(
  seq_len(nrow(pairs)),
  ~{
    base <- pairs$base_model[.x]
    img  <- pairs$image_model_col[.x]

    vec_noimg <- dat[[base]]
    vec_img   <- dat[[img]]

    test <- wilcox.test(vec_noimg, vec_img, exact = FALSE)

    broom::tidy(test) %>%
      mutate(model_base = base)
  }
)

mw_results
```

### 5.1 Multiple-Comparison Correction

```{r p-adjust}
mw_results_adj <- mw_results %>%
  mutate(
    p_adj_bonferroni = p.adjust(p.value, method = "bonferroni"),
    p_adj_fdr        = p.adjust(p.value, method = "fdr")
  ) %>%
  arrange(p_adj_bonferroni)

mw_results_adj
```

## 6. Effect Sizes (Rank-Biserial)

```{r effect-size}
eff_results <- purrr::map_df(
  seq_len(nrow(pairs)),
  ~{
    base <- pairs$base_model[.x]
    img  <- pairs$image_model_col[.x]

    vec_noimg <- dat[[base]]
    vec_img   <- dat[[img]]

    tmp <- tibble(
      score   = c(vec_noimg, vec_img),
      variant = c(
        rep("no_images", length(vec_noimg)),
        rep("with_images", length(vec_img))
      )
    )

    rstatix::wilcox_effsize(tmp, score ~ variant) %>%
      mutate(model_base = base)
  }
)

eff_results
```

## 7. Combined Summary Table

```{r combined-summary}
# Objective differences: mean & median differences (with_images - no_images)
diff_summary <- purrr::map_df(
  seq_len(nrow(pairs)),
  ~{
    base <- pairs$base_model[.x]
    img  <- pairs$image_model_col[.x]

    vec_noimg <- dat[[base]]
    vec_img   <- dat[[img]]

    tibble(
      model_base      = base,
      mean_no_images  = mean(vec_noimg, na.rm = TRUE),
      mean_with_images = mean(vec_img, na.rm = TRUE),
      median_no_images  = median(vec_noimg, na.rm = TRUE),
      median_with_images = median(vec_img, na.rm = TRUE),
      mean_diff       = mean(vec_img, na.rm = TRUE) - mean(vec_noimg, na.rm = TRUE),   # with - without
      median_diff     = median(vec_img, na.rm = TRUE) - median(vec_noimg, na.rm = TRUE)
    )
  }
)

summary_table <- mw_results_adj %>%
  select(model_base, statistic, p.value, p_adj_bonferroni, p_adj_fdr) %>%
  left_join(eff_results %>% select(model_base, effsize, magnitude), by = "model_base") %>%
  left_join(diff_summary, by = "model_base") %>%
  arrange(p_adj_bonferroni)

summary_table
```

## 8. Significance & Direction

```{r significance}
sig_table <- summary_table %>%
  mutate(
    significance = case_when(
      p_adj_bonferroni < 0.001 ~ "***",
      p_adj_bonferroni < 0.01  ~ "**",
      p_adj_bonferroni < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    direction = case_when(
      mean_diff >  0 ~ "with_images > no_images",
      mean_diff <  0 ~ "no_images > with_images",
      TRUE           ~ "no difference"
    )
  )

sig_table
```

## 9. Effect Size vs Adjusted p-value

```{r effect-vs-p, fig.width=7, fig.height=7}
sig_table %>%
  ggplot(aes(
    x = effsize,
    y = -log10(p_adj_bonferroni),
    label = model_base
  )) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 20) +
  geom_hline(yintercept = 0, linetype = "dashed") +   # horizontal at 0 (optional)
  geom_vline(xintercept = 0, linetype = "dashed") +   # vertical at 0 effect size
  labs(
    title = "Effect Size vs Adjusted p-value (With Images vs Without)",
    x = "Rank-biserial effect size (with vs without images)",
    y = "-log10(Bonferroni-adjusted p-value)"
  ) +
  theme_minimal()

```
