---
title: "Reasoning-Level Score Distributions"
author: "Anonymous"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(broom)
library(rstatix)
library(ggrepel)
```

## 1. Data Import

```{r load-data}
llm_file <- "../output_csvs/openai_reasoning_comparison_grade_dist.csv"

llms_raw <- readr::read_csv(llm_file, show_col_types = FALSE)

glimpse(llms_raw)
```

Assumptions based on the files:

* `llms_raw` has:

  * Column 1: index (0, 1, 2, …) → **ignore**
  * Columns 2+: one column per LLM reasoning variant → **model score distributions**

```{r clean-data}
variant_cols <- names(llms_raw)[-1]
llms <- llms_raw %>% select(all_of(variant_cols))

dim(llms)
variant_cols
```

## 2. Long Format for Plotting

We build a long data frame with:

* `model_base` – same-model family
* `reasoning_level` – reasoning setting
* `variant` – original column name
* `score` – numeric score

```{r comparison-pairs}
reasoning_levels <- c("med", "low", "minimal")

parse_reasoning_variant <- function(variant) {
  level <- stringr::str_extract(variant, "(high|med|low|minimal)$")
  if (is.na(level)) {
    return(tibble(variant = variant, model_base = NA_character_, reasoning_level = NA_character_))
  }
  tibble(
    variant = variant,
    model_base = stringr::str_remove(variant, paste0("_", level, "$")),
    reasoning_level = level
  )
}

variant_meta <- purrr::map_dfr(variant_cols, parse_reasoning_variant) %>%
  filter(!is.na(reasoning_level))

comparison_pairs <- variant_meta %>%
  filter(reasoning_level == "high") %>%
  transmute(model_base, high_variant = variant) %>%
  tidyr::crossing(comparison_level = reasoning_levels) %>%
  left_join(
    variant_meta %>%
      filter(reasoning_level %in% reasoning_levels) %>%
      transmute(model_base, comparison_level = reasoning_level, comparison_variant = variant),
    by = c("model_base", "comparison_level")
  ) %>%
  filter(!is.na(comparison_variant)) %>%
  mutate(
    comparison = paste0("high_vs_", recode(comparison_level, med = "medium")),
    pair_label = paste0(model_base, "_", comparison)
  )

comparison_pairs
```

```{r long-format}
llms_long <- llms %>%
  pivot_longer(
    cols = everything(),
    names_to = "variant",
    values_to = "score"
  ) %>%
  left_join(variant_meta, by = "variant")

llms_long %>%
  filter(!is.na(reasoning_level)) %>%
  group_by(model_base, reasoning_level, variant) %>%
  summarise(
    n    = n(),
    mean = mean(score, na.rm = TRUE),
    sd   = sd(score, na.rm = TRUE),
    .groups = "drop"
  )
```

## 3. Distribution Visualizations

### 3.1 Density Plot by Reasoning Level

```{r density-plot, fig.width=10, fig.height=6}
llms_long %>%
  filter(!is.na(reasoning_level), model_base %in% unique(comparison_pairs$model_base)) %>%
  ggplot(aes(x = score, fill = reasoning_level)) +
  geom_density(alpha = 0.4) +
  facet_wrap(vars(model_base)) +
  labs(
    title = "Score Distributions by Reasoning Level",
    x = "Score",
    y = "Density",
    fill = "Reasoning Level"
  ) +
  theme_minimal()
```

### 3.2 Boxplots by Reasoning Level

```{r boxplot-all, fig.width=10}
llms_long %>%
  filter(!is.na(reasoning_level), model_base %in% unique(comparison_pairs$model_base)) %>%
  mutate(reasoning_level = factor(reasoning_level, levels = c("high", "med", "low", "minimal"))) %>%
  ggplot(aes(x = reasoning_level, y = score)) +
  geom_boxplot(outlier.alpha = 0.4) +
  facet_wrap(vars(model_base)) +
  labs(
    title = "Score Distributions by Reasoning Level",
    x = "Reasoning Level",
    y = "Score"
  ) +
  theme_minimal()
```

## 4. Mann-Whitney U Tests (High vs Other Reasoning Levels)

We compare each model's **high reasoning** score distribution against the same model's **medium**, **low**, and **minimal** reasoning score distributions when both variants exist.

```{r mann-whitney}
rank_biserial <- function(x, y, statistic) {
  nx <- length(x)
  ny <- length(y)
  u <- as.numeric(statistic)
  (2 * u / (nx * ny)) - 1
}

mw_results <- purrr::map_df(
  seq_len(nrow(comparison_pairs)),
  function(i) {
    pair <- comparison_pairs[i, ]
    high_vec <- llms[[pair$high_variant]] %>% purrr::discard(is.na) %>% unlist()
    comparison_vec <- llms[[pair$comparison_variant]] %>% purrr::discard(is.na) %>% unlist()

    if (length(high_vec) == 0 || length(comparison_vec) == 0) {
      return(tibble())
    }

    test <- wilcox.test(high_vec, comparison_vec, exact = FALSE)

    broom::tidy(test) %>%
      mutate(
        model_base = pair$model_base,
        comparison = pair$comparison,
        high_variant = pair$high_variant,
        comparison_variant = pair$comparison_variant,
        n_high = length(high_vec),
        n_comparison = length(comparison_vec),
        high_mean = mean(high_vec),
        comparison_mean = mean(comparison_vec),
        mean_difference = high_mean - comparison_mean,
        effsize = rank_biserial(high_vec, comparison_vec, statistic)
      )
  }
)

mw_results
```

### 4.1 Multiple-Comparison Correction

We adjust p-values across all models using Bonferroni and FDR.

```{r p-adjust}
mw_results_adj <- mw_results %>%
  mutate(
    p_adj_bonferroni = p.adjust(p.value, method = "bonferroni"),
    p_adj_fdr        = p.adjust(p.value, method = "fdr")
  ) %>%
  arrange(model_base, comparison)

mw_results_adj
```

### 4.3 Combined Summary Table

```{r summary-table}
summary_table <- mw_results_adj %>%
  mutate(
    magnitude = case_when(
      abs(effsize) < 0.1 ~ "negligible",
      abs(effsize) < 0.3 ~ "small",
      abs(effsize) < 0.5 ~ "medium",
      TRUE ~ "large"
    )
  ) %>%
  select(
    model_base, comparison, high_variant, comparison_variant, n_high, n_comparison,
    statistic, p.value, p_adj_bonferroni, p_adj_fdr, effsize, magnitude,
    high_mean, comparison_mean, mean_difference
  ) %>%
  arrange(model_base, comparison)

summary_table
```

## 5. Significance Labels

```{r significance-flags}
sig_table <- summary_table %>%
  mutate(
    significance = case_when(
      p_adj_bonferroni < 0.001 ~ "***",
      p_adj_bonferroni < 0.01  ~ "**",
      p_adj_bonferroni < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )
# save as csv
write_csv(sig_table, "../statistical_results/human_vs_reasoning_model_significance_summary.csv")
sig_table
```

## 6. Effect Size vs Adjusted p-value

```{r effect-vs-p, fig.width=7, fig.height=7}
sig_table %>%
  mutate(label = paste(model_base, comparison, sep = ": ")) %>%
  ggplot(aes(
    x = effsize,
    y = -log10(p_adj_bonferroni),
    label = label
  )) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 20) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Effect Size vs Adjusted p-value (High vs Lower Reasoning)",
    x = "Rank-biserial effect size (positive = high reasoning higher)",
    y = "-log10(Bonferroni-adjusted p-value)"
  ) +
  theme_minimal()
```

```{r}
mean_diffs <- summary_table %>%
  select(model_base, comparison, high_mean, comparison_mean, mean_difference)

mean_diffs

```



```{r}
perf_pct <- mean_diffs %>%
  mutate(comparison_percent_of_high = 100 * comparison_mean / high_mean)

perf_pct

```


```{r}
final_ranked <- summary_table %>%
  left_join(perf_pct %>% select(model_base, comparison, comparison_percent_of_high),
            by = c("model_base", "comparison")) %>%
  arrange(p_adj_bonferroni)

final_ranked
```
