---
title: "Human vs LLM Score Distributions"
author: "Anonymous"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(broom)
library(rstatix)
library(ggrepel)
```

## 1. Data Import

```{r load-data}
human_file <- "../output_csvs/h_grade_distribution.csv"   # EDIT IF NEEDED
llm_file   <- "../output_csvs/openai_reasoning_comparison_grade_dist.csv"   # EDIT IF NEEDED

humans_raw <- readr::read_csv(human_file, show_col_types = FALSE)
llms_raw   <- readr::read_csv(llm_file, show_col_types = FALSE)

glimpse(humans_raw)
glimpse(llms_raw)
```

Assumptions based on the files:

* `humans_raw` has:

  * Column 1: index (0, 1, 2, …) → **ignore**
  * Column 2: `humans` → **human score distribution**
* `llms_raw` has:

  * Column 1: index (0, 1, 2, …) → **ignore**
  * Columns 2+: one column per LLM model → **model score distributions**

```{r clean-data}
# Human scores: use second column only (full distribution)
human_score_col <- names(humans_raw)[2]

humans <- humans_raw[[human_score_col]]

length(humans)

# LLMs: drop the first index column, keep the rest as models
model_cols <- names(llms_raw)[-1]
llms <- llms_raw %>% select(all_of(model_cols))

dim(llms)
model_cols[1:5]
```

## 2. Long Format for Plotting

We build a long data frame with:

* `model` – `"Human"` or LLM name
* `score` – numeric score

```{r long-format}
humans_long <- tibble(
  model = "Human",
  score = humans
)

llms_long <- llms %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "score"
  )

scores_long <- bind_rows(humans_long, llms_long)

scores_long %>%
  group_by(model) %>%
  summarise(
    n    = n(),
    mean = mean(score, na.rm = TRUE),
    sd   = sd(score, na.rm = TRUE),
    .groups = "drop"
  )
```

## 3. Distribution Visualizations

### 3.1 Density Plot for Humans vs a Few Models

```{r density-plot, fig.width=10, fig.height=6}
# Pick a few models to highlight (first 3, plus humans)
highlight_models <- c("Human", head(model_cols, 3))

scores_long %>%
  filter(model %in% highlight_models) %>%
  ggplot(aes(x = score, fill = model)) +
  geom_density(alpha = 0.4) +
  labs(
    title = "Score Distributions: Humans vs Selected Models",
    x = "Score",
    y = "Density"
  ) +
  theme_minimal()
```

### 3.2 Boxplots for Humans and All Models

```{r boxplot-all, fig.width=10}
scores_long %>%
  mutate(model = fct_reorder(model, score, .fun = median, na.rm = TRUE)) %>%
  ggplot(aes(x = score, y = model)) +
  geom_boxplot(outlier.alpha = 0.4) +
  labs(
    title = "Score Distributions: Human vs All LLM Models",
    x = "Score",
    y = "Model"
  ) +
  theme_minimal()
```

## 4. Mann–Whitney U Tests (Human vs Each Model)

We compare the **overall human score distribution** vs **each model’s score distribution** using the Wilcoxon rank-sum (Mann–Whitney U) test.

```{r mann-whitney}
human_vec <- humans

mw_results <- purrr::map_df(
  model_cols,
  ~{
    model_name <- .x
    model_vec  <- llms[[model_name]]

    test <- wilcox.test(human_vec, model_vec, exact = FALSE)

    broom::tidy(test) %>%
      mutate(model = model_name)
  }
)

mw_results
```

### 4.1 Multiple-Comparison Correction

We adjust p-values across all models using Bonferroni and FDR.

```{r p-adjust}
mw_results_adj <- mw_results %>%
  mutate(
    p_adj_bonferroni = p.adjust(p.value, method = "bonferroni"),
    p_adj_fdr        = p.adjust(p.value, method = "fdr")
  ) %>%
  arrange(p_adj_bonferroni)

mw_results_adj
```

### 4.2 Effect Size (Rank-Biserial Correlation)

```{r effect-size}
effect_sizes <- purrr::map_df(
  model_cols,
  ~{
    model_name <- .x
    model_vec  <- llms[[model_name]]

    tmp <- tibble(
      score = c(human_vec, model_vec),
      group = c(rep("Human", length(human_vec)),
                rep(model_name, length(model_vec)))
    )

    rstatix::wilcox_effsize(tmp, score ~ group) %>%
      mutate(model = model_name)
  }
)

effect_sizes
```

### 4.3 Combined Summary Table

```{r summary-table}
summary_table <- mw_results_adj %>%
  select(model, statistic, p.value, p_adj_bonferroni, p_adj_fdr) %>%
  left_join(
    effect_sizes %>%
      select(model, effsize, magnitude),
    by = "model"
  ) %>%
  arrange(p_adj_bonferroni)

summary_table
```

## 5. Significance Labels

```{r significance-flags}
sig_table <- summary_table %>%
  mutate(
    significance = case_when(
      p_adj_bonferroni < 0.001 ~ "***",
      p_adj_bonferroni < 0.01  ~ "**",
      p_adj_bonferroni < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )
# save as csv
write_csv(sig_table, "../statistical_results/human_vs_reasoning_model_significance_summary.csv")
sig_table
```

## 6. Effect Size vs Adjusted p-value

```{r effect-vs-p, fig.width=7, fig.height=7}
sig_table %>%
  ggplot(aes(
    x = effsize,
    y = -log10(p_adj_bonferroni),
    label = model
  )) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 20) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Effect Size vs Adjusted p-value (Human vs Model)",
    x = "Rank-biserial effect size",
    y = "-log10(Bonferroni-adjusted p-value)"
  ) +
  theme_minimal()
```

```{r}
mean_diffs <- scores_long %>%
  group_by(model) %>%
  summarise(mean_score = mean(score)) %>%
  ungroup() %>%
  mutate(
    human_mean = mean_score[model == "Human"],
    diff_vs_humans = mean_score - human_mean,
  )

mean_diffs

```



```{r}
perf_pct <- mean_diffs %>%
  mutate(percent_of_human = 100 * mean_score / human_mean)

perf_pct

```


```{r}
final_ranked <- summary_table %>%
  left_join(mean_diffs %>% select(model, diff_vs_humans), by="model") %>%
  left_join(perf_pct %>% select(model, percent_of_human), by="model") %>%
  arrange(desc(percent_of_human))

final_ranked
```
