---
title: "Single vs Multi-Context Comparison (sc vs mc)"
author: "Anonymous"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(broom)
library(rstatix)
library(ggrepel)
```

## 1. Data Import

We use the file that contains **only model variants** with `_sc` and `_mc` suffixes.

```{r load-data}
file_sc_mc <- "../output_csvs/single_vs_multiple_context_grade_dist.csv"  # EDIT PATH IF NEEDED

raw_sc_mc <- readr::read_csv(file_sc_mc, show_col_types = FALSE)

glimpse(raw_sc_mc)
```

Assumptions (based on the file):

* Column 1 (`Unnamed: 0`) is an **index** → we ignore it.
* Remaining columns are model variants of the form:

  * `modelname_sc`
  * `modelname_mc`

```{r clean-data}
# Drop the first index column
dat <- raw_sc_mc[, -1]

# Identify sc and mc columns
sc_cols <- names(dat)[str_ends(names(dat), "_sc")]
mc_cols <- names(dat)[str_ends(names(dat), "_mc")]

sc_cols
mc_cols
```

We now extract **base model names** so we can pair `_sc` and `_mc` columns:

```{r model-names}
# Helper to strip suffix
strip_suffix <- function(x, suffix) sub(paste0(suffix, "$"), "", x)

models_sc <- strip_suffix(sc_cols, "_sc")
models_mc <- strip_suffix(mc_cols, "_mc")

# Models that appear in both sc and mc
common_models <- intersect(models_sc, models_mc)
common_models
```

## 2. Long Format for sc vs mc

We create a long table with columns:

* `model`   – base model name (e.g., `"sonnet"`, `"o3"`)
* `context` – `"sc"` or `"mc"`
* `item_id` – row index (shared for sc & mc)
* `score`   – numeric score

```{r long-format}
# Build long data for all models
long_list <- lapply(common_models, function(m) {
  sc_name <- paste0(m, "_sc")
  mc_name <- paste0(m, "_mc")

  tibble(
    item_id = seq_len(nrow(dat)),
    model   = m,
    sc      = dat[[sc_name]],
    mc      = dat[[mc_name]]
  )
})

wide_sc_mc <- bind_rows(long_list)

# Convert to long context format
scores_long <- wide_sc_mc %>%
  pivot_longer(
    cols = c(sc, mc),
    names_to = "context",
    values_to = "score"
  )

scores_long
```

Quick summary: mean and SD in each condition for each model.

```{r summary-descriptives}
summary_stats <- scores_long %>%
  group_by(model, context) %>%
  summarise(
    n    = n(),
    mean = mean(score, na.rm = TRUE),
    sd   = sd(score, na.rm = TRUE),
    median = median(score, na.rm = TRUE),
    .groups = "drop"
  )

summary_stats
```

## 3. Visualizing sc vs mc

### 3.1 Boxplots per Model

```{r boxplots, fig.width=10, fig.height=0.4 * length(common_models) + 4}
scores_long %>%
  mutate(model = fct_reorder(model, score, .fun = median, na.rm = TRUE)) %>%
  ggplot(aes(x = context, y = score, fill = context)) +
  geom_boxplot(outlier.alpha = 0.4) +
  facet_wrap(~ model, scales = "free_y") +
  labs(
    title = "Single vs Multi-Context Scores (sc vs mc)",
    x = "Context",
    y = "Score"
  ) +
  theme_minimal()
```

### 3.2 Paired Differences (mc - sc)

We can also look at the **paired difference** per item:

```{r paired-differences, fig.width=10, fig.height=0.4 * length(common_models) + 4}
diff_long <- wide_sc_mc %>%
  mutate(diff = mc - sc)

diff_long %>%
  ggplot(aes(x = model, y = diff)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(outlier.alpha = 0.4) +
  coord_flip() +
  labs(
    title = "Paired Difference (mc - sc) per Model",
    x = "Model",
    y = "Difference in score (mc - sc)"
  ) +
  theme_minimal()
```

Positive diff = multi-context > single-context on average.
Negative diff = single-context > multi-context.

## 4. Paired Tests: sc vs mc within Each Model

Since scores are **paired by item** for sc and mc, we use the **Wilcoxon signed-rank test** (non-parametric paired test).

```{r paired-wilcoxon}
paired_results <- purrr::map_df(
  common_models,
  ~{
    m <- .
    subdat <- wide_sc_mc %>% filter(model == m)

    # Paired Wilcoxon signed-rank test: is sc distribution different from mc?
    test <- wilcox.test(subdat$sc, subdat$mc, paired = TRUE, exact = FALSE)

    broom::tidy(test) %>%
      mutate(model = m)
  }
)

paired_results
```

### 4.1 Multiple-Comparison Correction

We have one test per model; adjust p-values (Bonferroni + FDR):

```{r p-adjust}
paired_results_adj <- paired_results %>%
  mutate(
    p_adj_bonferroni = p.adjust(p.value, method = "bonferroni"),
    p_adj_fdr        = p.adjust(p.value, method = "fdr")
  ) %>%
  arrange(p_adj_bonferroni)

paired_results_adj
```

### 4.2 Effect Sizes for Paired sc vs mc

We use **rank-biserial effect size** for each model’s sc vs mc difference:

```{r effect-sizes}
eff_results <- purrr::map_df(
  common_models,
  ~{
    m <- .
    subdat <- scores_long %>%
      filter(model == m)

    # Paired effect size: sc vs mc
    rstatix::wilcox_effsize(
      data = subdat,
      score ~ context,
      paired = TRUE
    ) %>%
      mutate(model = m)
  }
)

eff_results
```

### 4.3 Combined Summary

We now combine:

* medians in each condition
* mean paired difference (mc - sc)
* Wilcoxon stats
* adjusted p-values
* effect sizes

```{r combined-summary}
# Compute mean/median per context + mean diff
diff_summary <- wide_sc_mc %>%
  group_by(model) %>%
  summarise(
    n_pairs   = n(),
    mean_sc   = mean(sc, na.rm = TRUE),
    mean_mc   = mean(mc, na.rm = TRUE),
    median_sc = median(sc, na.rm = TRUE),
    median_mc = median(mc, na.rm = TRUE),
    mean_diff = mean(mc - sc, na.rm = TRUE),   # mc - sc
    .groups   = "drop"
  )

summary_table <- paired_results_adj %>%
  select(model, statistic, p.value, p_adj_bonferroni, p_adj_fdr) %>%
  left_join(eff_results %>% select(model, effsize, magnitude), by = "model") %>%
  left_join(diff_summary, by = "model") %>%
  arrange(p_adj_bonferroni)

summary_table
```

## 5. Significance & Direction

Add a simple significance label and sign of effect:

```{r significance-labels}
sig_table <- summary_table %>%
  mutate(
    significance = case_when(
      p_adj_bonferroni < 0.001 ~ "***",
      p_adj_bonferroni < 0.01  ~ "**",
      p_adj_bonferroni < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    direction = case_when(
      mean_diff >  0 ~ "mc > sc",
      mean_diff <  0 ~ "sc > mc",
      TRUE           ~ "no difference"
    )
  )

sig_table
```

## 6. Visual: Effect Size vs Adjusted p-value

```{r effect-vs-p, fig.width=7, fig.height=7}
sig_table %>%
  ggplot(aes(
    x = effsize,
    y = -log10(p_adj_bonferroni),
    label = model
  )) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 20) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Paired sc vs mc: Effect Size vs Adjusted p-value",
    x = "Rank-biserial effect size (mc vs sc)",
    y = "-log10(Bonferroni-adjusted p-value)"
  ) +
  theme_minimal()
```
